---
title: "Math 152 - Statistical Theory - Homework 3"
author: "Ethan Ashby"
date: "Due: Friday, September 11, midnight PDT"
output: pdf_document
---


```{r warning=FALSE, comment=FALSE, message=FALSE, echo = FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center")
library(mosaic)
```


#### 8: R Baseball & Bayes

Consider the baselball problem we discussed in class.  (See website for a copy of the handout.)  Let $\alpha_0$ and $\beta_0$ be your prior parameters.

  a. What are your choices of $\alpha_0$ and $\beta_0$?  [Use $\alpha_0$ and $\beta_0$ values that were **not** discussed in class.] What features of the plot of the prior density function made you think these were good choices?    I wrote a function below that you can use to try out different options for the prior.

```{r}
betaplot<-function(a,b){
ex<-a/(a+b)
varx<-a*b/((a+b)^2*(a+b+1))
plotDist('beta', params = list(a,b),
         main = paste("alpha=",a,", beta=",b),
         key=list(space="right",
         text=list(c(paste("E=",round(ex,4)),
                      paste("Var=",round(varx,4)),
		paste("SD=", round(sqrt(varx),4)), paste("P(>=0.2)=", round(pbeta(0.2, a, b, lower.tail=FALSE), 4))))))
}
betaplot(a=2.5,b=30) # change a and b to something we did not do in class
```

I choose $\alpha_0=2.5$ and $\beta_0=30$ for it's center and spread. I'm skeptical that this player is capable of performing at a major league level. This distributions' center and (prior expected batting average=0.08) and probability of exceeded a barely adequate level (P(batting at 0.2)=0.0175) jive with my prior belief that this player would not succeed in the majors.

  b. Using properties of expectation [that is, consider both estimates as functions of *X*, not of $\theta$], find the bias and variance of $\hat{\theta}_f$ and $\hat{\theta}_b$.  You are a frequentist here, and your answers should both be functions of $\theta$.
  
  (i)
  $$MSE(\hat{\theta}_f)=E[(\hat{\theta}_f-\theta)^2]=Var(\hat{\theta}_f)+[E(\hat{\theta}_f)-\theta]^2$$
  $$E(\hat{\theta}_f)=\frac{1}{n}E(x)=\frac{1}{n}(n\theta)=\theta$$
  $$Var(E(\hat{\theta}_f))=\frac{1}{n^2}Var(X)=\frac{1}{n^2}n\theta(1-\theta)=\frac{\theta(1-\theta)}{n}$$
  $$\textrm{Bias}: [E(\hat{\theta}_f)-\theta]^2=[\theta-\theta]^2=0$$
  $$\textrm{Variance}: Var(E(\hat{\theta}_f))=\frac{\theta(1-\theta)}{10}$$
  
  (ii)
  **Note that $X$ denotes the number of successful trials: i.e. base hits**
  $$\hat{\theta}_b=\frac{X+\alpha}{n+\alpha+\beta}$$
  $$E(\hat{\theta}_b)=\frac{E(X)+\alpha}{n+\alpha+\beta}=\frac{n\theta+\alpha}{n+\alpha+\beta}$$
  $$Var(E(\hat{\theta}_b))=Var(\frac{X+\alpha}{n+\alpha+\beta})=\frac{Var(X)}{(n+\alpha+\beta)^2}=\frac{n\theta(1-\theta)}{(n+\alpha+\beta)^2}$$
  $$\textrm{Bias}: [E(\hat{\theta}_b)-\theta]^2=\frac{n\theta+\alpha}{n+\alpha+\beta}-\theta=\frac{-(\alpha+ \beta)\theta+\alpha}{n+\alpha+\beta}=\frac{-32.5 \theta +2.5}{42.5}$$
  $$\textrm{Variance}: Var(E(\hat{\theta}_b))=\frac{n\theta(1-\theta)}{(n+\alpha+\beta)^2}=\frac{10\theta-10\theta^2}{42.5^2}$$
  
  c. Based on your comparison of the MSE, do you recommend using $\hat{\theta}_f$ or $\hat{\theta}_b$? Explain.  [Hint: first determine whether one estimator has a smaller MSE.  Over what region?]

I would use the bayesian estimator in this case. The bayesian estimator has a lower MSE over the range $\theta \in [0.02, 0.25]$ which seems like a reasonable range of batting average for this not super accomplished batter.

```{r fig.height=5,  fig.width=6}
theta<-seq(0,1,.01) # theata is his true batting average
plot(theta, theta*(1-theta)/10,type="l",lty=1,
     xlab="true batting average",ylab="MSE") #mse.f
lines(theta, (10*theta-10*theta^2)/42.5^2+((-32.5*theta+2.5)/42.5)^2,lty=2) #mse.b
abline(v=0.02)
abline(v=0.25)
rect(0.02, -0.1, 0.25, 0.025, col="red", density = 7, border = "red", lty=3)
legend(x="topright",c("Frequentist MSE", "Bayesian MSE @ your parameters"),
		lty=c(1:5), bg="white", cex=0.5)
mtext("MSE for different estimators of batting average",line=1,cex=1)
```

  d. If John Spurrier gets three hits in ten at bats, what is your estimate of $\theta$?  (Given your answer to c.)
  
  $\hat{\theta}_b=\frac{X+\alpha}{n+\alpha+\beta}=\frac{3+2.5}{10+2.5+30}=0.129$

  e. Show that in the beta-binomial family, $\hat{\theta}_b$ is a weighted average of $\hat{\theta}_f$ and the prior mean.
  
  $$\hat{\theta}_b=\frac{X+\alpha}{n+\alpha+\beta}$$
  $$\hat{\theta}_f=\frac{X}{n}$$
  $$E[\xi(\theta)]=\frac{\alpha}{\alpha+\beta}$$
  
  $$\hat{\theta}_b=\frac{n}{n+\alpha+\beta}*\frac{X}{n}+\frac{\alpha+\beta}{n+\alpha+\beta}*\frac{\alpha}{\alpha+\beta}=\gamma(\frac{X}{n})+(1-\gamma)(\frac{\alpha}{\alpha+\beta})$$
  
  This is a weighted average of the prior mean and frequentist estimator.  
  
  
  
  
